Supplementary Materials

### 2023-06-23 manuscript/supplementary.tex
ms updates

>>>>

\section{Theory}
\subsection{Overview of Quantitative Genetic Models of Effective Population Size}

Here, we step through a quick derivation of \textcite{Santiago1995-hx}.
Throughout, we assume random mating, hermaphroditic individuals, and a constant
population size. The change in a neutral allele's frequency in one generation
can be partitioned into the three sources of stochasticity: the random
associations with fitness backgrounds (i.e. \emph{draft}), the non-heritable
randomness in family size, and the Mendelian noise from heterozygotes
segregating. If we let $x_{0,i} \in \{0, \nicefrac{1}{2}, 1\}$ be the frequency
of neutral alleles individual $i$ in generation 0 carries, we can partition the
random neutral allele frequency of the population ($x_t$ without the individual
index) into the underlying stochastic causes,

\begin{align}
  x_1 = \frac{1}{2N} \sum_{i=1}^N \left( x_{0,i}k_{0,i} + \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \right)
\end{align}

where $k_{0,i}$ is the number of surviving gametes parent $i$ passes on, and
$\delta_{0,i,j}$ is a random term that encapsulates the noise due to Mendelian
segregation of heterozygotes. If parent $i$ is a homozygote ($x_{0,i} \in \{0,
1\}$), then $\delta_{0,i,j} = 0$, whereas if $x_{0,i} = \nicefrac{1}{2}$ then
$\delta_{0,i,j} = \pm \nicefrac{1}{2}$ with equal probability. This is because
a heterozygous parent will transmit half a neutral allele in expectation, but
each round of Mendelian segregation must pass on either 0 or 1 alleles, which
the random $\pm \nicefrac{1}{2}$ term imposes. The factor of $\nicefrac{1}{2}$
is due to the fact that we're summing over $N$ diploids, but considering the
number of gametes they transmit. Since each diploid parent must have two
offspring to maintain a constant population size, $\nicefrac{1}{N} \sum_i
k_{0,i} = 2$.

The frequency in the initial generation is $x_0 = \nicefrac{1}{N} \sum_{i=1}^N
x_{0,i}$, though to indicate that we treat this as fixed, we use $p_0 = x_0$.
Then, the allele frequency change is,

\begin{align}
  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \\
  \Delta x_1 &= K_1 + H_1
\end{align}

where $K_1$ and $H_1$ are the random change in neutral allele frequency change
due to offspring number (including heritable and non-heritable components), and
Mendelian segregation.

Now, let us look the variance of $\var(\Delta x_1)$ over evolutionary
replicates. Since the Mendelian random process and the offspring random process
are independent,

\begin{align}
  \var(\Delta x_1) &= \var(K_1) + \var(H_1).
\end{align}

Looking at each term,

\begin{align}
  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var\left(x_{0,i} (k_{0,i} - 1) \right)
\end{align}

where we ignore the covariance terms due to the sum, since these are on order
$\nicefrac{1}{N^3}$. In the first generation from an arbitrary starting point,
the neutral alleles assort independently into diploids with respect to their
fitness, so

\begin{align}
  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var(x_{0,i}) \var(k_{0,i} - 1)  \\
            &\approx \frac{1}{4N} \var(x_{0,i}) \var(k_{0,i}).
\end{align}

Assuming that no correlation between parental gametes (e.g. no inbreeding),
$\var(x_{0,i})$ is the binomial variance in individual allele frequency, or
$p_0(1-p_0)/2$, and $\var(k) := \var(k_{0,i})$ is the offspring variance of an
individual given by the reproduction process. For example, if the reproduction
process is a neutral multinomial Wright--Fisher, $\var(k) \approx 2$.

\begin{align}
  \var(K_1) &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4}.
\end{align}

The Mendelian noise variance term can be derived similarly. First, the sum over
each parent's transmitted gametes can be simplified by noting that parents are
exchangeable over evolutionary replicates with respect to their contribution to
this term. In a constant population size, the double summation over parents and
their offspring can be replaced by a summation over offspring, since both sum
$N$ exchangeable terms. Then, note that $\var(\delta_{i,j}) = \nicefrac{1}{4}
p_0(1-p_0)$, so

\begin{align}
  \var(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
\end{align}

Finally, we have

\begin{align}
  \var(\Delta x_1) &= \var(K_1) + \var(H_1) \\
  &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \\
                   &\approx \frac{p_0(1-p_0)}{2N}\left(\frac{\var(k)}{4} + \frac{1}{2}\right)
\end{align}

(c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).
Note that if the reproduction process is a neutral multinomial Wright--Fisher,
$\var(k) \approx 2$, and this simplifies to the expected Wright--Fisher
variance. If we were to \emph{define} a variance effective population size by
setting this variance in neutral allele frequency change to the expected
variance under a Wright--Fisher model ($V_\text{WF}$), we'd have

\begin{align}
  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \\
  V_\text{WF} &= \var(\Delta p_1) \\
  N_e &= \frac{4N}{\var(k) + 2} \\
\end{align}

(c.f. \cite{Wright1938-tv}).

Now, we look at this variance equation behaves with $t > 1$.

\begin{align}
  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
\end{align}

and

\begin{align}
  \label{eq:var_wf}
  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + \mathcal{C}_{1,2} \\
          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + \mathcal{C}_{1,2}
\end{align}

Consequently, the variance of the neutral allele frequency changes each
generation according to the probability of failing to coalesce each generation
and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up due to associations
with heritable fitness backgrounds. Note that the covariance term $\mathcal{C}_{1,2} =
\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] = 0$ under neutral
evolution, and this simplifies to the well-known equation for variance in a
Wright--Fisher population.

Now, we turn our attention to the case where there is heritable fitness
variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
terms emerge when $\var(k)$ has a heritable component of fitness that can be
transmitted along with the neutral allele, thereby affecting the neutral
allele's trajectory in later generations. If we partition the offspring number
into heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
$\var(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
individuals, $V_h > 0$. Because the population size is assumed to be constant,
the mean heritable fitness across individuals is constrained to be $\E_i(f_i) =
1$; this implies that the individual fitness values $f_i$ are relative
fitnesses as is standard in population genetics, in this case to the population
mean. Thus $V_h$ is the squared coefficient of heritable fitness variation
(this is often denoted as $C^2 = \nicefrac{V_A}{\bar{w}^2}$, see
\cite{Crow1958-pc,Charlesworth1987-ab,Houle1992-ur}).

We can further decompose the $K_1$ term into $K_1 = S_1 + D_1$,

\begin{align}
  \var(\Delta x_1) &= \var(S_1) + \var(D_1) + \var(H_1) \\
  &\approx \frac{p_0(1-p_0)}{2N} V_h + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2}. \\
\end{align}

(c.f. \cite{Santiago1995-hx} equation 11).

To see how the covariance terms accumulate, consider $\var(x_3 - p_0)$. Note
that $\E(S_t) = \E(D_t) = \E(H_3) = 0$, since which neutral allele we track is
arbitrary, so by symmetry the expected change is zero. Then,

\begin{align}
  \label{eq:var_x}
  \var(x_3 - p_0) &= \E\left[ \left(S_1 + D_1 + H_1 + S_2 + D_2 + H_2 + S_3 + D_3 + H_3 \right)^2 \right] \\
                  &= \E({S_1}^2) + \E({S_2}^2) + \E({S_3}^2)  \\
                  &+ \E(S_1 S_2) + \E(S_1 S_3) + \E(S_2 S_3) \\
                  &+ \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) + \E({D_3}^2) + \E({H_3}^2)
\end{align}

The covariance terms $\E(S_i S_j)$ for $j > i$ represent the expected neutral
allele frequency change from a neutral allele becoming associated with a
fitness background in generation $i$ and that fitness association persisting
until generation $j$. However, the covariances terms $\E(S_j S_i)$ for $j > i$
are zero since associations in the future cannot affect the past (see p. 1041
of \cite{Buffalo2019-qs}). Let us consider $S_1$ and $S_2$. We have,

\begin{align}
  S_1 = \frac{1}{2N} \sum_{i=1}^N x_{0,i}(f_{0,i} - 1) = \cov(x_{0,i}, f_{0,i}) \\
  S_2 = \frac{1}{2N} \sum_{i=1}^N x_{1,i}(f_{1,i} - 1) = \cov(x_{1,i}, f_{0,i})
\end{align}

which are chance covariances (across the population, \emph{not} evolutionary
replicates) created by the random sorting of neutral alleles and fitness into
individuals, \emph{each generation}. These covariances are equivalent to the
Robertson-Price equation \parencite{Robertson1966-fs,Price1970-si}, which
predict the change in neutral frequency due to heritable fitness (and likewise
with the change due to non-heritable factors).

Considering the underlying haplotypes that lead to $x_{0,i}$ and $f_{0,i}$ can
give us an equation for the dynamics of the $\E(S_i S_j)$ (for $j > i$) terms
over evolutionary replicates. Let us partition the allele frequency and fitness
by into the average of the paternal contributions. The neutral allele frequency
per gamete is either 0 or 1, and we assume the fitnesses across gametes are
additive. Then, as long as there is random mating, we can simplify the
associations in a diploid by looking at a single gamete,

\begin{align}
  S_1 = \cov(x_{0,i}, f_{0,i}) &= \cov\left(\frac{x_{0,i}' + x_{0,i}''}{2}, f_{0,i}' + f_{0,i}''\right)  \\
                               &= \frac{1}{2}\left(\cov(x_{0,i}', f_{0,i}' + f_{0,i}'') + \cov(x_{0,i}'', f_{0,i}' + f_{0,i}'')\right)  \\
                               &= \cov(x_{0,i}', f_{0,i}' + f_{0,i}'')  \\
                               &= \cov(x_{0,i}', f_{0,i}') + \cov(x_{0,i}', f_{0,i}'')  \\
                               &= S_1' + S_1''
\end{align}

which follows from symmetry, since it does not matter which gamete carrying the
neutral allele we track (c.f. \cite{Santiago1998-bs} p. 2107). The primes now
indicate whether the selected locus is on the same gamete as the neutral
allele (single prime), or the homologous gamete (double prime).

After meiosis, a fraction $1-r$ of these associations between the tracked
neutral neutral allele $x_{0,1}'$ and fitness background $f_{0,i}'$ persist,
and a fraction $r$ disassociate its currently coupled background and create an
association with the homologous fitness background, $f_{0,i}''$. When linkage
is tight, this latter term can be ignored as we will do here, though it does
impact background levels of neutral diversity \parencite{Santiago1995-hx}.
Additionally, the fitness effects of the selected background may change in
later generations, in complex ways \parencite{Barton1986-yh,Turelli1990-kd}.
\textcite{Santiago1995-hx} assume an equilibrium level of fitness variation
$V_h$, with the fitness variance associated with a particular background
decaying at a simple geometric decay at rate $1-\kappa$ each generation, which
is sufficient for background selection. An approximate rate of the decay can be
worked out for the particular selective system. In general, this can be a much
more complicated function; \textcite{Buffalo2019-qs} show that even fluctuating
selection can be accommodated.

Now, the associations created in generation 1 and that persist in future
generations as (ignoring $S_i''$ terms) follow the pattern

\begin{align}
  S_{2}' &= S_1 (1-\kappa)(1-r) \\
  S_{3}' &= S_1 (1-\kappa)^2(1-r)^2 \\
  S_{4}' &= S_1 (1-\kappa)^3(1-r)^3.
\end{align}

Then, the cumulative effect of the fitness associations created in generation
$1$ can be written as,

\begin{align}
  \label{eq:qterm}
  S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
  Q_t &= 1 + \sum_{i=1}^t (1-r)^i(1-\kappa)^i
\end{align}

or for continuous $t$,

\begin{align}
  \label{eq:q_t}
  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
\end{align}
which as $t \to \infty$, converges to,

\begin{align}
  \label{eq:Qinf}
  Q_\infty &= \frac{1}{\kappa + r(1-\kappa)}.
\end{align}

This shows the long reach of heritable fitness factors in altering allele
frequency through the generations. In terms like $\E(S_1 S_2)$, a fraction
$S_1'$ of the heritable allele frequency change $S_2$ were \emph{previously}
built associations between the neutral allele and a fitness background from
$S_1$. As long as the direction of selection is the same, the expected change
of this fraction of associations formed in the first generation to later
generations 2 and 3 would be,

\begin{align}
  \E(S_1 S_2) &= \E(S_1^2) (1-\kappa)(1-r) = \E(S_1^2) Q_2^2 \\
  \E(S_1 S_3) &= \E(S_1^2) (1-\kappa)^2(1-r)^2 = \E(S_1^2) Q_3^2.
\end{align}

With these covariance terms, we now have a full model for $\var(p_t - p_0)$. We
will look at the case for $t=2$. As shown in equation XXX, the second moment of
every term is a function of the variance in neutral allele frequency in that
generation, $x_t(1-x_t)$. This variance does not stay its initial value of
$x_0(1-x_0)$; it decays each generation due to coalesce from the drift and
selection processes modeled by this approach. The rate at which this decay
happens is the effective population size each generation implied by this model. 
We can write,

\begin{align}
    \label{eq:vardecay}
  \var(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \\
                  & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \\
  \frac{\var(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_h}{2N_{e,1}} + \frac{V_h Q_2^2}{2N_{e,2}}\left(1-\frac{1}{2N_{e,1}}\right)  +  \frac{V_h}{2N_{e,2}}  \left(1-\frac{1}{2N_{e,1}}\right)  \\ 
                                     &+ \frac{V_n}{8N_{e,1}} + \frac{1}{4N_{e,1}} + \frac{V_n}{8N_{e,2}}  \left(1 - \frac{1}{2N_{e,1}}\right) 
+ \frac{1}{4N_{e,2}} \left(1 - \frac{1}{2N_{e,1}}\right).
\end{align}

In general, the evolution of the variance in allele frequency change in a
system with heritable fitness due to a single locus $r$ recombination fraction
away from the neutral site is given by,

\begin{align}
  \label{eqn:var_xt}
  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^\infty \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{t=1}^{i-1} \left(1-\frac{1}{2 N_{e,t}}\right). \\
\end{align}

Since the probability of coalesce (or, equivalently, identity by descent) is
proportional to the variance in allele frequency change
\parencite{Barton2000-zg}, this encodes the pairwise coalescent rate of the
population through time under selection through the $N_{e,t}$ terms. Others
have found that the genealogies under purifying selection can be characterized
by such a time-dependent effective population size
\parencite{Nicolaisen2013-gv}. In the original derivation of
\textcite{Santiago1995-hx}, they set each $N_{e,t}$ term to a single $N$ and
solve this summation in the limit to infinity giving,

\begin{align}
  \label{eq:var_asymp}
  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \frac{4 V_h Q_\infty^2 +V_n+2}{4 (2 N-1)}.
\end{align}

To relate this to an asymptotic effective population size, note that in an
ideal Wright--Fisher population,

\begin{align}
  \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_e}\right)^t \right]
\end{align}

such that if we look at the change in the variance in allele frequency change
between adjacent generations $t$ and $t-1$, we get an implied $N_{e,t}$

\begin{align}
  \label{eqn:ne_var}
  N_{e,t} = \frac{p_0(1-p_0) - \var(x_{t-1})}{2(\var(x_t) - \var(x_{t-1}))}
\end{align}
%
(cf. p. 1018, \cite{Santiago1995-hx}).

Solving this for equation \eqref{eq:var_asymp} gives the asymptotic effective
population size,

\begin{align}
  %N_e &= \frac{2 (2 N-1)}{4 V_h Q_\infty^2 +V_n+2} \\
  N_e &\approx \frac{4N}{4 V_h Q_\infty^2 +V_n+2}
\end{align}

(cf. equation 18, \cite{Santiago1995-hx}). Levels of pairwise diversity,
however, depend on the full distribution of pairwise coalesce times, or the
$N_{e,t}$. While this would require solving the recursion defined by equations
\eqref{eqn:var_xt} and \eqref{eqn:ne_var}, \textcite{Santiago1998-bs} show that
a reasonably good approximation follows from using a single effective
population size for all timepoints, $N_{e,t} = N_e$, but use the time-indexed
$Q_t$. This leads to,

\begin{align}
  \label{eq:ne_t}
  N_{e,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
\end{align}

(c.f. equation 15 of \cite{Santiago1995-hx} and p. 1270 of \cite{Santiago2016-mu}).

\subsection{Effective Population Size Under Polygenic Selection}

The results of the previous section assume only a single locus a recombination
fraction $r$ apart from the neutral locus is determining the coalesce rate. The
model of \textcite{Santiago1998-bs} extends this to the case where fitness is
polygenic. As they show in their original paper, this system can accommodate a
variety of different equilibrium selection systems as long as expressions for
$V_h$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.

First, we extend equation \eqref{eq:ne_t} to a polygenic system. Throughout, a
multiplicative fitness model is assumed (i.e. independent effects, no
epistasis), where the fitness of individual $i$ is the product of their fitness
contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^n
w_{i,l}$. Then, assuming mean fitness is one and independence between sites,
the total fitness variation is multiplicative across the locus-level fitnesses,

\begin{align}
  V_h %&= \E_i\left( w_i^2 \right) - 1 \\
      %&= \prod_{l=1}^n (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
      &= \prod_{l=1}^n (1 + V_{h,l})  - 1
\end{align}

Similarly, the $Q_t^2 V_h$ term in the single-locus model is multiplicative
under the polygenic model, as the change in site-specific fitness variances are
assumed independent,

\begin{align}
  Q_t^2 V_h &= \prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
\end{align}

where $Q_{t,l}$ is the cumulative impact of selection due to the selected site
$l$ (c.f. equation 4 \cite{Santiago1998-bs}). The factor of $\nicefrac{1}{2}$
is due to the fact that we are ignoring the chance that the fitness background
on the homologous chromosome recombines onto the haplotype with the tracked
neutral allele (i.e. due to the $Q_t''$ associations). In other words, only
half the fitness variation in a diploid can stay associated with the neutral
allele through the generations.

To simplify the derivation, assume $V_n = 2$ as under a Wright--Fisher model.
Then, the effective population size is,

\begin{align}
  N_{e,t} &\approx \frac{N}{V_h Q_t^2 + 1} \\
          &\approx \frac{N}{\prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \\
          &\approx N \exp\left(- \sum_{l=1}^n \frac{V_{h,l} Q_{t,l}^2}{2}\right)
\end{align}

(c.f. \cite{Santiago1998-bs} equation 4b). If we write each site $l$'s
contribution to the variance as a deviation from the average contribution,
$V_{h,l}^2 = \nicefrac{1}{n} V_h + \varepsilon_l$, then

\begin{align}
  N_{e,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \varepsilon_l\right)  Q_{t,l}^2\right) \\
          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \varepsilon_l  Q_{t,l}^2\right)
\end{align}

Santiago and Caballero assume that over evolutionary replicates, as selected
sites are randomly distributed, there is not systematic covariance between the
fitness variation at a site and the decay of its association with the neutral
site, e.g. $\E(\varepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
term.

\begin{align}
      N_{e,t} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2\right) \\
\end{align}

\subsection{Continuous Approximation to the Segment Under Selection}

Our model differs from that of \textcite{Santiago1998-bs} in that we do not
integrate over the entire genome. In their 1998 model, Santiago and Caballero
consider the impact of both sites under selection on the same chromosome, as
well as sites on independently assorting chromosomes. We only consider the
contribution of a single segment under purifying selection, and the reduction
from additional segments accumulate multiplicatively (similar to our
implementation of the classic BGS model, XXX). Like their model, we imagine the
reduction experienced by a focal neutral site in the middle of the segment.
The implementation of our method differs slightly on though, see XXX.

Since $Q_{t,l}$ terms depend on $l$ only through the recombination fraction
between the neutral allele and the selected site, $r(l)$. Assuming a constant
per-basepair recombination rate of $r_\text{BP}$, $r(l) = r_\text{BP} l$ we can
approximate the sum with an integral for the asymptotic $Q_\infty$. The
reduction experienced in the middle of a chromosome is twice the reduction
experienced from each half,

\begin{align}
      N_{e,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \\
              &\approx N \exp\left(- \frac{V_{h}}{M} \int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr \right) \\
\end{align}

where $M = n r_\text{BP}$ is the total segment length in Morgans. Then, the
total asymptotic effect of associations from all sites as $Q_\infty^2$.  Then,

\begin{align}
  Q_\infty^2 &= \frac{2}{M}\int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr  \\
  &=  \frac{2}{M} \int_0^{\nicefrac{M}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr  \\
  &= \frac{2}{(1-Z) (2-(2-M) Z)}
\end{align}

Santiago and Caballero further approximate this; we can derive their
approximation by setting $Z = 1-\kappa$ and doing a first-order Taylor series
expansion around $\kappa$ (since the loss in variance is presumed to be small).
Then,

\begin{align}
  Q_\infty^2 = \frac{2}{M (1-Z)} + \frac{2 (M-2)}{M^2}+\frac{2 \left(M^2-4 M+4\right) (1-Z)}{M^3} + \mathcal{O}(\kappa^2)
\end{align}

Santiago and Caballero's approximation only keeps this first term, e.g. their
$Q_\infty^2 \approx 2/(M(1-Z))$, which is only accurate in the domain $M >
0.2$.

\begin{align}
  N_{e,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)M}\right) 
\end{align}

(c.f. \cite{Santiago1998-bs} equation 8). For small $M$, the small $\kappa$
approximation strongly deviates from the full equation; thus, throughout, we
use the unapproximated version,

\begin{align}
  \label{eq:Ne_inf}
  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
\end{align}

For the reasons described earlier, a single asymptotic $N_{e,\infty}$ cannot
fully characterize pairwise coalescent rates in all cases, and thus neutral
diversity levels depend on the full $N_{e,t}$, for which we integrate equation
\eqref{eq:q_t} to get,

\begin{align}
  \label{eq:net_integral}
  N_{e,t} \approx N \exp\left(-\frac{V_h}{M} \int_0^{\nicefrac{M}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
\end{align}

\subsection{The Variance Dynamics}

The heritable variance in fitness $V_h$ associated with the tracked neutral
allele does not remain constant through the generations, but decays due to
selection and drift, and is maintained by the input of new mutations. For an
arbitrary polygenic selective system, the dynamics are extraordinarily
complicated due to the complexity of multilocus selection, and the dynamics of
the trait response to selection depend on the variance, which in turn depends
on higher moments of the fitness distribution. We discuss the simplification of
Santiago and Caballero's model here; hereafter, we assume that $V_h = V_A$, the
genetic variance in additive fitness put on a relative scale such that
$\E_i(f_i) = 1$.

Each generation, the genetic variance changes due to changes in the allele
frequencies at selected sites and linkage disequilibria between these sites.
These are typically expressed as recursions, but certain models permit these
changes to be expressed as a proportional reduction (e.g. truncation selection
and stabilizing selection models; \cite{Keightley1988-eq} p. 36, and
\cite{Walsh2018-bt} p. 557). Santiago and Caballero's model follows this
approach, assuming the variance decays due to selection and drift at a constant
rate $Z := 1-\kappa$ through the generations, such that the variance in the next
generation excluding mutation is $V_A' = (1-\kappa) V_A$ and $\Delta V_A =
-\kappa V_A$.

With mutation, $\Delta V_A = V_m - \kappa V_A$. The model of Santiago and
Caballero assumes that under the long-run equilibrium, $\Delta V_A = 0$, such
that $\kappa V_A = V_m$, and thus the rate of loss is $\kappa =
\nicefrac{V_m}{V_A}$. Then the variance for the dynamics simplify to,

\begin{align}
  \label{eq:Z}
  V_A' &= \left(1 - \frac{V_m}{V_A}\right) V_A \\
  Z &:= \frac{V_A'}{V_A} = 1 - \kappa = 1 - \frac{V_m}{V_A}
\end{align}

(c.f. \cite{Santiago1998-bs} equation 9).

% TODO this is a good description that could make its way to maintext

\subsection{Effective Population Size Under Strong Background Selection}

Under strong background selection, multiple segregating deleterious sites
contribute to fitness variation. BGS models assume a multiplicative fitness
model, such that the fitness of an individual is

\begin{align}
  w = (1-s)^{n}
\end{align}

where $n$ is the number of deleterious mutations this individual carries.

Classic background selection theory assumes an infinite population size, such
that mutation frequencies are at their mutation-selection equilibrium and the
expected number of mutations across individuals is $\E(n) = \nicefrac{\mu}{s}L$
where $L$ is the total number of basepairs that can be mutated (the mutational
target size).  Under a multiplicative fitness function in an infinite
population, selection cannot generate linkage disequilibria
\parencite{Turelli1990-kd}. Thus the classic strong background selection model
assumes additive genetic fitness is equal to the additive genic fitness (which
excludes the contribution of covariance between selected sites),

\begin{align}
  \label{eq:va_bgs}
  V_\text{BGS} &:= \V(w) = 2 s^2 \sum_{l=1}^L p_l(1-p_l)  \\
            &= 2 L s^2 \frac{\mu}{s} \left( 1 - \frac{\mu}{s}\right) = U s + \mathcal{O}(\mu^2) \approx Us.
\end{align}

since under mutation-selection balance $p_l = \nicefrac{\mu}{s}$ and $U$ is
defined $U := 2 L \mu$, the deleterious mutation rate per diploid genome, per
generation. They then set $V_A = V_\text{BGS}$, and note that $V_m \approx
Us^2$, since mutations have frequency of $\nicefrac{1}{2N}$ and each increase
the variance by a factor of $s^2$. Note that because the classic BGS model
assumes deterministic selection dynamics (i.e. an infinite population), the
equilibrium $Z$ given by equation \eqref{eq:Z} excludes the reduction due to
drift. The reduction factor under the strong BGS model is,

\begin{align}
  Z_\text{BGS} := 1 - \frac{Vm}{V_\text{BGS}} = 1 - \frac{Us^2}{Us} = 1 - s
\end{align}

(c.f. \cite{Santiago1998-bs} equation 10). An alternate derivation can be found
by working out the deterministic reduction in variance from the single-locus
dynamics.

Now, we can use these values for $Z_\text{BGS}$, $V_m$, and $V$ in equation
\eqref{eq:Ne_inf}. Through this alternate quantitative genetics derivation, we
arrive at the same equation as classic BGS theory,

\begin{align}
  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right).
\end{align}

This matches equation 8 of \textcite{Hudson1995-xc} except for the $Ms$ term
and equation 10 of \textcite{Nordborg1996-nq} except for the $2s$ term.

\subsection{Effective Population Size Under Weak and Strong BGS} \label{supp:weak-strong}

With the possibility of weak selection, deleterious alleles can drift up in
frequency and fix. Consequently, the fitness distribution is no longer
stationary, as the set of haplotypes without any deleterious mutations can be
lost due to drift. This changes the shape of the fitness distribution, and thus
the variance parameter that our model relies upon
\parencite{Gessler1995-hz,OFallon2010-my,Good2013-lp,Haigh1978-gt,Higgs1995-xc}.
So far, finding an explicit equation for the variance has been difficult due to
the ``moment-closure problem": the variance depends on higher moments such as
the skew, which in turn depends on higher moments. While the weak selection
regime is composed of an interaction of mutation, selection, drift, and linkage
processes and the dynamics that differ qualitatively based on parameter
combinations. In their approximation, \textcite{Santiago2016-mu} consider the
modified additive genic variance due to the fixation of weakly deleterious
mutations. This fixation process is related to the Muller's Ratchet, the rate
$R$ of fixation of deleterious mutations per generation per region, since each
fixation directly reduces the additive genic variance of the fitness
distribution. However, in general, the rate of this ratchet is unknown since it
depends on the fitness variance (and thus higher moments of the fitness
distribution).

% NOTE: The S&C math using Fisher's FToNS argues fitness decreases
% at a rate of -s/T each generation. However their model is haploid;
% in a diploid model, fixation makes everyone  a homozygote with fitness
% effect -2s; thus the rate must be -2s/T which agrees with sim

In their 2016 paper, Santiago and Caballero derive a general equation for the
fitness variance that accounts for the reduction due to the fixation of weakly
deleterious alleles (i.e. the rate of Muller's ratchet). There derivation for
the variance uses a heuristic approach based on Fisher's Fundamental Theorem of
Natural Selection (equation 2, \cite{Santiago2016-mu}, equation 15A
\cite{Garcia-Dorado2007-jj}). A more formal derivation follows from
\textcite{Higgs1995-xc} (though they assume a haploid asexual system, the end
result is the same up to a factor of two). In their work, they derive equations
for the change per generation for the moments and cross-moments of the fitness
distribution under multiplicative selection in a haploid model. They find, the
haploid rate of the ratchet is $r = u - s v_n$ (equation 13.3,
\cite{Higgs1995-xc}), where we use lowercase to distinguish between haploid and
diploid models. Here, $v_n$ is the variance in the \emph{number} of deleterious
mutations; as long as selection is not too strong, the haploid additive
variance is $v_{A} \approx s^2 v_n$. Substituting this for $v_n$ and
rearranging, the haploid fitness variance is related to the $v_A = us - rs$,
which is identical to Santiago and Caballero's (\citeyear{Santiago2016-mu})
equation 2, noting that the waiting time between fixations (i.e. ratchet
clicks) is $T = \nicefrac{1}{r}$. Intuitively, the fitness variance $Us$ is
reduced by the rate at which deleterious mutations fix and remove fitness
variation from the population.

Now, note that under a diploid model, the fitness variance is twice that as the
haploid model, so $V_A = 2 v_A$. Then, the fitness variance is,

\begin{align}
  \label{supp-eq:Va}
  V_{A} = Us - 2Rs. 
\end{align}

where $U = 2u$ and $R=r$ is the diploid rate of the ratchet, which depends on
higher moments of the fitness distribution, but only differs from $r$ through a
rescaling of the population size. Note that if selected sites are sufficiently
deleterious that they cannot fix (or the population is infinite), $R=0$ and the
additive fitness variance is identical to that under strong background
selection, $V_A = Us = V_\text{BGS}$ (equation \eqref{eq:va_bgs}). Then, the
fixation probability of a new deleterious mutation with additive effects ($1-s$
in the heterozygote, $1-2s$ in the homozygote) in a diploid population is,

\begin{align}
  p_\text{fix} = \frac{1-\exp(\nicefrac{2 N_e s}{N})}{1-\exp(4 N_e s)} \approx
  \frac{2 N_e s}{N \left(\exp(4 N_e s)-1\right)}
\end{align}

(c.f. \cite{Durrett2008-ql}, equation 7.21; \cite{Kimura1957-rk}, equation
5.6). Then, Santiago and Caballero argue that the rate of the ratchet is the
inverse of the averaging waiting time until a fixation, $T \approx 1/(UN
p_\text{fix})$ (note that since $U$ is the \emph{diploid} mutation rate, the
total population mutation rate is $UN$ per generation), so

\begin{align}
  T \approx \frac{\exp(4 N_e s)-1}{2 U s N_e}
\end{align}

Since the ratchet rate depends on $N_e$, which in turn depends on the rate of
the ratchet, we write the ratchet, or substitution, rate per basepair per
generation as,

\begin{align}
  R(N_e) := \frac{2 U s N_e}{\exp(4 N_e s)-1}.
\end{align}

Under weak and strong background selection the variance input of mutations
$V_m$ is the same, and using the $V_A$ derived above, the decay in variance is
